---
title: "Results"
author: "Isabelle van der Zanden and Gelieke Steeghs"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```

``` {r package}

library(Rmisc)
library(tidyverse)
library(kableExtra)
library(flextable)
library(ggpubr)
library(rstatix)
library(readxl)
library(writexl)
library(car)
library(wesanderson)
library(multcomp)
library(multcompView)
library(emmeans)
library(broom)
library(broom.mixed)
library(ggcorrplot)
library(reshape2)
library(lmerTest)
library(factoextra)
library(vegan)
library(GGally)
library(ggpmisc)
library(RColorBrewer)

```

``` {r upload and mutate files}
lml <- read_csv2("MASTER_LML.csv")

lml$litter_sp <- factor(lml$litter_sp, levels=c("Alder","Hazel","Chestnut"))
lml$mesh_size <- factor(lml$mesh_size, levels=c("Fine","Medium","Coarse"))
lml <- lml %>% mutate(litter_mass_loss_final = litter_mass_loss_final*100)

mt <- read_csv2("MASTERFILE_decomp_10.csv")
cnp <- read_csv("CNP_data.csv")

litter_cn <- read_csv2("litter_cn.csv")
litter_lignin <- read_csv2("litter_lignin.csv")
litter_lignin <- litter_lignin %>% dplyr::rename("lignin" = "lignin (mg/g)")

mt_1 <- mt %>% left_join(cnp, by="code")

mt_2 <- mt_1 %>% mutate(age = 2021-year_of_landuse,
                      earthworms_2021 = compost_worms + soil_eating_worms + pendelaars)

mt_2 <- mt_2 %>% rowwise() %>% dplyr::mutate(earthworms_2022 = sum(dplyr::c_across(contains("Abundance")))*10)
mt_2 <- mt_2 %>% dplyr::rename("vb_code" = "code")

# join litter mass data with landuse & location data
mt_2_lu_loc <- mt_2 %>% dplyr::select(vb_code, landuse, location)
lml_2 <- lml %>% left_join(mt_2_lu_loc, by ="vb_code")

# filter out only the 4 subplots that have been used for the litter mass experiment
# mt_3 <- mt_2 %>% filter(plot_nr < 5)

# sum columns mites and earthworms
mt_4 <- cbind(mt_2, mites_sum = rowSums(mt_2[c('Mites_oribatid_kg','Mites_other_kg')]))
mt_4 <- cbind(mt_4, biomass_ew_total = rowSums(mt_2[c('Biomass_anecic','Biomass_endogeic','Biomass_epigeic','Biomass_juvenile')])*10)

# sum bacteria + actinos
mt_4 <- cbind(mt_4, bact_total = rowSums(mt_4[c("Bacteria_total","Actinobacteria_total")]))

#loclu
mt_4$location <- factor(mt_4$location)
levels(mt_4$location) <- c("L4","L3","L1", "L2")

mt_4$loc_lu <- paste(mt_4$location, mt_4$landuse, sep="_")

# join litter mass data with physical, chemical, biological data
lml_mt <- lml_2 %>% left_join(mt_4, by ="vb_code")

```

# Results

## Soil characteristics  
When comparing soil characteristics between food forests and grasslands significant differences were found with higher moisture and soil organic matter content, lower bulk density, higher oribatid and other (mesostigmatic and prostigmatic) mite abundance for food forests, but significantly higher earthworm abundance and biomass and higher AMF biomass for grasslands. No significant differences were found for pH, nutrient concentrations (C, N, P), springtail abundance and fungal (non-AMF), bacterial, actinobacterial, general microbial biomass and total microbial biomass (table 1). 

<!-- Average AMF biomass in grassland was `r format(mt_4 %>% filter(landuse == "Grassland") %>% summarize(mean = mean(AMF_NLFA, na.rm=TRUE)), digits=2)` ug/gr. (example) -->

``` {r table}

# list of model outputs ANOVA
models <- list()
for (i in c('ph_kcl','som_percentage','b.d.','moisture_percentage', 'ec',
            'Total_P_mg_kg','Olsen_P_mg_kg','C_total','N_total', 
            'earthworms_2022','biomass_ew_total',
            'Springtails_nr_kg','Mites_oribatid_kg','Mites_other_kg',
            'bact_total','Fungi_total','AMF_NLFA')) {
  f <- formula(paste(i,"~ landuse + (1|location/landuse)"))
  models[[i]] <- lmer(f, data=mt_4, na.action = "na.omit")
}

tab_order <- c('b.d.','moisture_percentage','ph_kcl','som_percentage', 'ec',
               'Total_P_mg_kg', 'Olsen_P_mg_kg','C_total','N_total',
               'bact_total','AMF_NLFA','Fungi_total',              
               'Springtails_nr_kg','Mites_oribatid_kg','Mites_other_kg',
               'earthworms_2022','biomass_ew_total')

mt_wide <- mt_4 %>% pivot_longer(cols=c(ec,ph_kcl:b.d.,
                                        Total_P_mg_kg:Olsen_P_mg_kg,
                                        Fungi_total, AMF_NLFA:Springtails_nr_kg,
                                        C_total:N_total,
                                        earthworms_2022,biomass_ew_total,bact_total), names_to="measurement")
mt_wide <- mt_wide %>% mutate(category = if_else(measurement=='b.d.'|measurement=='moisture_percentage',
                                                 "Physical",
                                                 if_else(measurement=='ph_kcl'|measurement=='som_percentage'|measurement=='ec'|
                                                measurement=='Total_P_mg_kg'|measurement== 'Olsen_P_mg_kg'|measurement=="C_total"|measurement=="N_total",
                                              "Chemical","Biological")))
                                              # if_else(measurement=='Bacteria_total'|measurement=='Microbial'|
                                              #          measurement=='Actinobacteria_total'|
                                              #          measurement=='Fungi_total'|measurement=='Total_plfa'|
                                              #          measurement=='AMF_NLFA',"Microbial biomass",
                                              # if_else(measurement=='Springtails_nr_kg'|
                                              #          measurement=='Mites_oribatid_kg'|
                                              #          measurement=='Mites_other_kg',
                                              #        "Micro-arthropod","Earthworms")
                                              # ))))

tab <- mt_wide %>% group_by(landuse, category, measurement) %>%
  summarise(mean = sprintf("%1.2f", mean(value, na.rm = TRUE)),
            sd = sprintf("(%1.2f)", sd(value, na.rm = TRUE))) %>%
  unite("value",mean, sd, sep=" ") %>%
  arrange(factor(measurement, levels = tab_order)) %>%
  pivot_wider(names_from=landuse, values_from=value)

# #alternative way
# x <- mt_wide %>% group_by(landuse, category, measurement) %>% 
#   summarise(
#     across(.cols = 'value',
#            .fns = list(
#              mean = ~ mean(.x, na.rm = TRUE),
#              sd = ~ sd(.x, na.rm = TRUE)
#            )), .groups = "drop")

tab %>% 
  flextable() %>% 
  labelizor(labels = c(category = "Category", measurement = "Measurements", Food_forest = "Food forest", 
                       Total_P_mg_kg = "Total fosfor (mg/kg)", Olsen_P_mg_kg = "Olsen P (mg/kg)", 
                       b.d. = "Bulk density (g/cm3)", ph_kcl = "pH (KCl)", ec="Electrical conductivity (mS/m)",
                       C_total = "Total C (%)", N_total = "Total N (%)",
                       moisture_percentage = "Soil moisture (%)", 
                       som_percentage = "Soil organic matter (%)", 
                       earthworms_2022 = "Earthworms abundance (#/m2)",
                       biomass_ew_total = "Earthworms biomass (g/m2)",
                       Springtails_nr_kg = "Springtails abundance (#/kg)", 
                       Mites_oribatid_kg = "Mites oribatid abundance (#/kg)",
                       Mites_other_kg = "Mites other abundance (#/kg)",
                       bact_total = "Bacterial biomass (µg/g)",
                       Fungi_total = "Non-AMF fungal biomass (µg/g)",
                       AMF_NLFA = "AMF biomass (µg/g)")) %>%
  merge_v(j = ~ category) %>% 
  hline(j = ~ category, part = "all") %>%
  set_table_properties(layout = "autofit") %>%
  set_caption(caption = "Table 1. ANOVA results comparing food forests to grasslands for chemical, physical and biological variables. Table shows mean (standard deviation), F statistic and p-values.", 
                  style = "Table Caption")

```

## check earthworm correlations

``` {r scatterplot}
mt_4 %>% ggplot(aes(x=ec, y=earthworms_2022, fill=loc_lu)) +
  geom_point(size=3.5, pch=21) + scale_fill_brewer(palette = "Paired")
cor.test(mt_4$ec, mt_4$earthworms_2022, method="pearson", use = "complete.obs")

```

## Litter characteristics 

``` {r litter cn lignin table}

# For the litter characteristics, I used lm() instead of lmer() as I couldn't identify random effects?

# Litter CN data statistics
litter_cn %>% 
  group_by(litter_species) %>%
  summarize(
    mean1 = mean(c_proc),
    sd1 = sd(c_proc),
    mean2 = mean(n_proc),
    sd2 = sd(n_proc),
    mean3 = mean(cn_ratio),
    sd3 = sd(cn_ratio),
  )

cn.lm <- lm(cn_ratio ~ litter_species, data=litter_cn)
summary(cn.lm)
cn_anova <- anova(cn.lm)
cn_anova

cld(lsmeans(cn.lm,~ litter_species,adjust="tukey"), Letters = c(letters))

# Litter lignin data statistics

litter_lignin %>% 
  group_by(litter_species) %>%
  summarize(
    min = min(lignin),
    max = max(lignin),
    mean = mean(lignin),
    sd = sd(lignin)
  )

lignin.lm <- lm(lignin ~ litter_species, data=litter_lignin)
summary(lignin.lm)
lignin_anova <- anova(lignin.lm)
cld(lsmeans(lignin.lm,~ litter_species,adjust="tukey"), Letters = c(letters))

# C
c.lm <- lm(c_proc ~ litter_species, data=litter_cn)
hist(residuals(c.lm))
shapiro.test(residuals(c.lm))
anova(c.lm)
cld(lsmeans(c.lm,~ litter_species,adjust="tukey"), Letters = c(letters))

# N
n.lm <- lm(n_proc ~ litter_species, data=litter_cn)
hist(residuals(n.lm))
anova(n.lm)
cld(lsmeans(n.lm,~ litter_species,adjust="tukey"), Letters = c(letters))

```

## Litter mass loss data  
An interaction between land use, litter species and mesh size was observed for litter mass loss in coarse chestnut litter bags in grassland sites. Litter mass loss was lowest here. 
Litter mass loss was highest for hazel with medium and coarse mesh size and alder with the coarse mesh size. Medium levels of litter mass loss were found for hazel with fine mesh size and alder with fine and medium mesh size. Lowest levels of litter mass loss found for chestnut leaves for all mesh sizes. Effect of land use clear for alder with the coarse mesh size. Higher values were found in grasslands than food forests. 

``` {r statistics litter mass loss, fig.dim = c(10,8)}

# Descriptive statistics
# summarySE(lml_2, measurevar="litter_mass_loss_final", na.rm=TRUE) # for overall summary
# summarySE(lml_2, measurevar="litter_mass_loss_final", groupvars=c("landuse"), na.rm=TRUE)

# Levene's test for homogeneity of variances
lml_2 %>% group_by(landuse, location, litter_sp, mesh_size) %>%
          summarize(mean=mean(litter_mass_loss_final, na.rm=TRUE),
                    sd=sd(litter_mass_loss_final, na.rm=TRUE),
                    var=var(litter_mass_loss_final, na.rm=TRUE))
                    
variances_lml <- leveneTest(litter_mass_loss_final ~ landuse * location, data = lml_2)
variances_lml

# Histogram of the raw data
gghistogram(lml_2, x='litter_mass_loss_final', fill = 'lightgrey', na.rm=TRUE)

# Three-way ANOVA
threeway_slopes <- lmerTest::lmer(litter_mass_loss_final ~ landuse * litter_sp * mesh_size + (1 + location|landuse),data=lml_2, na.action = "na.omit")
threeway.lmeFit <- lmerTest::lmer(litter_mass_loss_final ~ landuse * litter_sp * mesh_size + (1|location/landuse),data=lml_2, na.action = "na.omit")
summary(threeway_slopes)
augmented_threeway <- augment(threeway.lmeFit)
ggplot(augmented_threeway, aes(x = .fitted, y = .resid)) + 
  geom_point()
outlierTest(threeway.lmeFit) # sample from line 115 is outlier?
threeway_anova <- anova(threeway.lmeFit)
anova(threeway.lmeFit,threeway_slopes)

# ANOVA results in table (still need to manually replace the significant and very small P-values by <0.0001)
litter_tab <- threeway_anova[,3:6]
litter_tab <- cbind(rownames(litter_tab), data.frame(litter_tab, row.names=NULL))
colnames(litter_tab)[1] <- "Parameter"

table_test <- litter_tab %>%
  flextable %>%
    labelizor(labels = c(landuse = "Land use", litter_sp = "Litter species", mesh_size = "Mesh size", 'landuse:litter_sp' = "Land use x Litter species", 'landuse:mesh_size' = "Land use x Mesh size", 'litter_sp:mesh_size' = "Litter species x Mesh size", 'landuse:litter_sp:mesh_size' = "Land use x Litter species x Mesh size", NumDF = "Num. DF", DenDF = "Den. DF", F.value = "F-value", Pr..F. = "P-value")) %>%
      colformat_double(j = c("DenDF"), digits = 0) %>%
        colformat_double(j = c("F.value", "Pr..F."), digits = 3) %>%
          set_table_properties(layout = "autofit") %>%
  set_caption(caption = "Table 2. ANOVA results comparing litter mass loss with an interaction effect between land use, litter species and mesh size, including degrees of freedom, F statistic and p-values.", 
                  style = "Table Caption")
table_test

# Three-way interaction letters are there but it looks kind of messy (probably better to do it manually again)
threeway_cld <- cld(lsmeans(threeway.lmeFit,~ landuse * litter_sp * mesh_size,adjust="tukey"), Letters = c(letters))
threeway_cld

# Testing the residuals
ggqqplot(residuals(threeway.lmeFit))
shapiro.test(residuals(threeway.lmeFit)) # it is significant...

# # Arcsine transformation (does not seem to make any difference)
# lml_3 <- lml_2
# lml_3$litter_mass_loss_final <- asin(sqrt(lml_2$litter_mass_loss_final))
# threeway_3.lmeFit <- lmer(litter_mass_loss_final ~ landuse * litter_sp * mesh_size + (1|location/landuse),data=lml_3, 
#                       na.action = "na.omit")
# ggqqplot(residuals(threeway_3.lmeFit))
# shapiro.test(residuals(threeway_3.lmeFit))

# Plot the outcomes
image <- lml_2 %>% group_by(mesh_size, litter_sp) %>%
  ggplot(aes(x=landuse, y=litter_mass_loss_final, fill=landuse)) +
  geom_boxplot(outlier.shape = NA, alpha=0.8, show.legend = FALSE) +
  scale_fill_manual(labels=c("Food forest","Grassland")) + 
  scale_fill_brewer(palette="Paired") +
  scale_fill_manual(values=c("#33A02B","#FF7F00")) +
  geom_point(position = position_jitterdodge(0.4), 
             pch=16, color="black",show.legend = FALSE) +
  geom_text(data = threeway_cld, aes(x = landuse, y = upper.CL, label = .group)) +
  facet_grid(litter_sp ~ mesh_size) +
  theme_bw(base_size = 20) +
  scale_x_discrete(labels=c("Food forest", "Grassland")) +
  labs(y="Litter mass loss (%)", x="Land use", fill="Landuse")
image
ggsave(file="3x3plot_flipped.svg", plot=image, width=10, height=10)


``` 

# Supplement

## Correlation matrix variables  

AMF positively correlated to pH, negatively to Olsen P, positively to springtails and negatively to oribatid mites. 

All PLFA measures roughly similarly correlated to other variables. Negatively to bulk density and positively to moisture and nutrients. Also positively to earthworms and nematodes.  

Earthworms positively correlated to almost everything except bulk density and mites and springtails.  

```{r correlation matrix, fig.dim=c(12,12)}
corr <- round(cor(mt_4 %>% dplyr::select(c('ph_kcl','som_percentage','b.d.','moisture_percentage', 'ec',
            'Total_P_mg_kg','Olsen_P_mg_kg','C_total','N_total',
            'Bacteria_total','Microbial','Actinobacteria_total','Fungi_total','Total_plfa','AMF_NLFA',
            'Springtails_nr_kg','Mites_oribatid_kg','Mites_other_kg',
            'earthworms_2022','biomass_ew_total')),
            use='pairwise.complete.obs'), 1)

# automatic ggcorrplot
# ggcorrplot(corr, method = "circle")

# Get lower triangle of the correlation matrix
  get_lower_tri<-function(corr){
    corr[upper.tri(corr)] <- NA
    return(corr)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(corr){
    corr[lower.tri(corr)]<- NA
    return(corr)
  }

lower_tri <- get_lower_tri(corr)

# more adjustable ggplot version
ggplot(melt(lower_tri, na.rm=TRUE), aes(Var1, Var2, fill=value)) +
  geom_tile(height=0.8, width=0.8) +
  scale_fill_gradient2(low="blue", mid="white", high="red", limits=c(-1,1)) +
  theme_minimal() +
  coord_equal() +
  labs(x="",y="",fill="Corr") +
  theme(axis.text.x=element_text(size=18, angle=45, vjust=1, hjust=1, 
                                 margin=margin(-3,0,0,0)),
        axis.text.y=element_text(size=18, margin=margin(0,-3,0,0)),
        panel.grid.major=element_blank(),
        legend.justification = c(1, 0),
        legend.position = c(0.6, 0.7),
        legend.direction = "horizontal") +
  geom_text(aes(Var1, Var2, label = value), color = "black", size = 4)

ggsave("corrplot.png", width=10, height=10)

# variables <- mt_4 %>% dplyr::select(c('landuse','ph_kcl','som_percentage','b.d.','moisture_percentage',
#             'Total_P_mg_kg','Olsen_P_mg_kg','C_total','N_total',
#             'Total_plfa','AMF_NLFA',
#             'Springtails_nr_kg','Mites_oribatid_kg','Mites_other_kg',
#             'earthworms_2022'))
# ggpairs(variables, ggplot2::aes(colour=landuse))


``` 

# PCA

``` {r PCA }
#pca with earthworms + micro-arthropods
mt_4_pca <- mt_4 %>% dplyr::select(c(landuse,location,loc_lu,ph_kcl,moisture_percentage,som_percentage,
                                     b.d.,ec,Total_P_mg_kg,Olsen_P_mg_kg,
                                     C_total,N_total,
                                     Total_plfa,AMF_NLFA,
                                     Mites_oribatid_kg,Mites_oribatid_kg,Mites_other_kg,
                                     Springtails_nr_kg,earthworms_2022,biomass_ew_total))

colSums(is.na(mt_4_pca))
mt_4_pca_nomiss <- na.omit(mt_4_pca) # 3 plots only for which there is earthworm and arthropod data
colSums(is.na(mt_4_pca_nomiss))

mt_4_pca_done <- mt_4_pca_nomiss %>% dplyr::select(!c(landuse,location,loc_lu))

pca <- prcomp(mt_4_pca_done, scale=TRUE)
# fviz_eig(pca) #check scree plot

# fviz_pca_ind(pca,
#              col.ind = "cos2", # Color by the quality of representation
#              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#              repel = TRUE,     # Avoid text overlapping
#              mean.point = FALSE # Remove mean point
#              )

# fviz_pca_ind(pca, label="none", habillage=mt_4_pca_nomiss$location,
#             addEllipses=TRUE, ellipse.level=0.95, mean.point = FALSE)

fviz_pca_biplot(pca, label="var", palette = "Paired", 
                pointshape=21, fill.ind = mt_4_pca_nomiss$loc_lu, 
                col.ind = "black",
                pointsize=4, mean.point = FALSE, col.var = "#767676",
                legend.title="Locations and land uses", 
                repel=TRUE, title="")+ 
                xlim(-4,4) + ylim(-4,4) +
                theme(legend.position="bottom",
                      text = element_text(size = 14)) +
                guides(fill=guide_legend(title.position="top", 
                                     title.hjust =0.5)) + 
                xlab("PC1 (41.5%)") +
                ylab("PC2 (20.7%)")


ggsave("pca_small.png", width=8, height=8)
ggsave("pca_small.svg")

``` 

``` {r PCA 2}

# PCA without earthworms + micro-arthropods

mt_4_pca_big <- mt_4 %>% dplyr::select(c(ph_kcl,moisture_percentage,som_percentage, b.d.,ec,
                                     Total_P_mg_kg,Olsen_P_mg_kg,C_total,N_total,bact_total,Fungi_total,AMF_NLFA
                                     ))
colSums(is.na(mt_4_pca_big))


pca_2 <- prcomp(mt_4_pca_big, scale=TRUE)
# fviz_pca_ind(pca_2,
#              col.ind = "cos2", # Color by the quality of representation
#              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#              repel = TRUE     # Avoid text overlapping
#              )

# fviz_pca_ind(pca_2, label="none", habillage=mt_4$loc_lu,
#              addEllipses=TRUE, ellipse.level=0.95, , palette = "Paired", pointshape=16, mean.point = FALSE)
# 
# fviz_pca_ind(pca_2, label="none", habillage=mt_4$loc_lu, palette = "Paired", pointshape=16, 
#              mean.point = FALSE)

fviz_pca_biplot(pca_2, label="var", palette = "Paired",
                pointshape=21, fill.ind = mt_4$loc_lu, 
                col.ind = "black",
                pointsize=4, mean.point = FALSE, col.var = "#767676",
                legend.title="Locations and land uses", 
                repel=TRUE, title="") + 
                xlim(-4,4) + ylim(-4,4) +
                theme(legend.position="bottom",
                      text = element_text(size = 14)) +
                guides(fill=guide_legend(title.position="top", 
                                     title.hjust =0.5)) + 
                xlab("PC1 (55.2%)") +
                ylab("PC2 (24.5%)")

ggsave("pca_big.png", width=8, height=8)
ggsave("pca_big.svg")

ind <- get_pca_ind(pca_2)

# ind$coord
``` 

# PERMANOVA?

``` {r adonis}
char <- mt_4 %>% dplyr::select(c('ph_kcl','som_percentage','b.d.','ec','moisture_percentage',
            'Total_P_mg_kg','Olsen_P_mg_kg','C_total','N_total',
            'bact_total','Fungi_total','AMF_NLFA'))

char_dist <- vegdist(char, method="bray")

char_div <- adonis2(char ~ landuse*location, data=mt_4,permutations = 999, method="bray")

char_div

char_tab <- tidy(char_div) %>% flextable()
char_tab

```


#multiple regression

``` {r multiple regression}
#Alder

mr_alder <- lml %>% filter(litter_sp == "Alder" & mesh_size == "Coarse") %>%
                  left_join(mt_4, by = "vb_code")

# mult_regr_alder <- lm(litter_mass_loss_final ~ ph_kcl + b.d. + NH4_mg_kg + Olsen_P_mg_kg + biomass_ew_total + Springtails_nr_kg + mites_sum + Total_plfa + AMF_NLFA, data = mr_alder)
# 
# summary(mult_regr_alder)
# 
# stepAIC(mult_regr_alder, direction="both")

#Hazel
mr_hazel <- lml %>% filter(litter_sp == "Hazel" & mesh_size == "Coarse") %>%
                  left_join(mt_4, by = "vb_code")

# mult_regr_hazel <- lm(litter_mass_loss_final ~ ph_kcl + b.d. + NH4_mg_kg + Olsen_P_mg_kg + biomass_ew_total + Springtails_nr_kg + mites_sum + Total_plfa + AMF_NLFA, data = mr_hazel)
# 
# summary(mult_regr_hazel)
# 
# stepAIC(mult_regr_hazel, direction="both")

#Chestnut
mr_chestnut <- lml %>% filter(litter_sp == "Chestnut" & mesh_size == "Coarse") %>%
                  left_join(mt_4, by = "vb_code")

# mult_regr_chestnut <- lm(litter_mass_loss_final ~ ph_kcl + b.d. + NH4_mg_kg + Olsen_P_mg_kg + biomass_ew_total + Springtails_nr_kg + mites_sum + Total_plfa + AMF_NLFA, data = mr_chestnut)
# 
# summary(mult_regr_chestnut)
# 
# 
# stepAIC(mult_regr_chestnut, direction="both")

``` 


``` {r pca including litter mass loss data}
#PCA with litter mass loss coarse mesh size alder and earthworms and micro-arthropods
mr_alder$loc_lu <-paste(mr_alder$location, mr_alder$landuse, sep="_")

mr_alder_pca <- mr_alder %>% dplyr::select(c(landuse,location,loc_lu, litter_mass_loss_final, ph_kcl,moisture_percentage,som_percentage, b.d., NO2_NO3_mg_kg,
                                     NH4_mg_kg,Total_P_mg_kg,Olsen_P_mg_kg,Total_plfa,AMF_NLFA,
                                     Mites_oribatid_kg,Mites_oribatid_kg,Mites_other_kg,
                                     Springtails_nr_kg,earthworms_2022,biomass_ew_total))

colSums(is.na(mr_alder_pca))
mr_alder_pca_nomiss <- na.omit(mr_alder_pca)
colSums(is.na(mr_alder_pca_nomiss))

mr_alder_pca_done <- mr_alder_pca_nomiss %>% dplyr::select(!c(landuse,location,loc_lu))

pca_3 <- prcomp(mr_alder_pca_done, scale=TRUE)
fviz_eig(pca_3)

fviz_pca_ind(pca_3, label="none", habillage=mr_alder_pca_nomiss$loc_lu, palette = "Paired", pointshape=16, mean.point = FALSE)  + xlim(-4,4) + ylim(-4,4) 

fviz_pca_biplot(pca_3, label="var", habillage=mr_alder_pca_nomiss$loc_lu, palette = "Paired", pointshape=16, mean.point = FALSE)  + xlim(-4,4) + ylim(-4,4) + theme(legend.position="bottom", ,text = element_text(size = 17))

ggsave("pca_smallest.png", width=8, height=8)


#PCA with litter mass loss coarse mesh size alder WITHOUT earthworms and micro-arthropods
mr_alder_pca_wo <- mr_alder %>% dplyr::select(c(vb_code,landuse,location,loc_lu, litter_mass_loss_final, ph_kcl,moisture_percentage,som_percentage, b.d., 
                                                NO2_NO3_mg_kg,NH4_mg_kg,Total_P_mg_kg,Olsen_P_mg_kg,Total_plfa,AMF_NLFA))
mr_alder_pca_wo <- mr_alder_pca_wo %>% arrange(vb_code)


colSums(is.na(mr_alder_pca_wo))
mr_alder_pca_wo_nomiss <- na.omit(mr_alder_pca_wo)
colSums(is.na(mr_alder_pca_wo_nomiss))

mr_alder_pca_done_wo <- mr_alder_pca_wo_nomiss %>% dplyr::select(!c(vb_code,landuse,location,loc_lu))

pca_4 <- prcomp(mr_alder_pca_done_wo, scale=TRUE)
fviz_eig(pca_4)


fviz_pca_ind(pca_4, label="none", habillage=mr_alder_pca_wo_nomiss$loc_lu, palette = "Paired", pointshape=16, mean.point = FALSE)  + xlim(-4,4) + ylim(-4,4) + theme(legend.position="bottom",text = element_text(size = 17))

fviz_pca_biplot(pca_4, label="var", habillage=mr_alder_pca_wo_nomiss$loc_lu, palette = "Paired", pointshape=16, mean.point = FALSE)  + xlim(-4,4) + ylim(-4,4) + theme(legend.position="bottom",text = element_text(size = 17))

ggsave("pca_lml.png", width=8, height=8)

ind_4 <- get_pca_ind(pca_4)
PC1 <- ind_4$coord[,1]
PC2 <- ind_4$coord[,2]
``` 

# Microbial and faunal effects

``` {r faunal effect}
lml_wide <- lml_2 %>% dplyr::select(-c("ff_code","treatment_code","litter_mass_loss","litter_mass_loss_G","notes")) %>% pivot_wider(names_from = c(mesh_size, litter_sp), values_from = litter_mass_loss_final)

lml_wide <- lml_wide %>% arrange(vb_code)

lml_wide <- lml_wide %>% mutate(alder_macro = Coarse_Alder-Medium_Alder,
                    hazel_macro = Coarse_Hazel-Medium_Hazel,
                    chestnut_macro = Coarse_Chestnut-Medium_Chestnut,
                    alder_meso = Medium_Alder-Fine_Alder,
                    hazel_meso = Medium_Hazel-Fine_Hazel,
                    chestnut_meso = Medium_Chestnut-Fine_Chestnut)

# filter out only the 4 subplots that have been used for the litter mass experiment
mt_5 <- mt_4 %>% filter(plot_nr < 5)

lml_wide <- left_join(lml_wide, mt_5, by = "vb_code")

lml_wide <- lml_wide %>% mutate(sum_arthropods = Mites_oribatid_kg + Mites_other_kg + Springtails_nr_kg,
                    sum_bact_fun = bact_total + Fungi_total)

# PCA only litter mass loss samples
# lml_wide <- lml_wide %>% filter(!grepl("VB002|VB116|VB118",vb_code)) #only use this if you want to plot the faunal effects against PC1/PC2

# lml_wide_long <- lml_wide %>% pivot_longer(cols = c("alder_macro","hazel_macro","chestnut_macro","alder_meso","hazel_meso","chestnut_meso"), values_to="faunal_effect")

lml_wide %>% ggplot(aes(x=sum_bact_fun, y=Fine_Alder)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p

lml_wide %>% ggplot(aes(x=sum_bact_fun)) + 
  geom_point(aes(y=Fine_Alder), color="#66c2a5", size=3) + stat_smooth(aes(y=Fine_Alder),method="lm", se=FALSE, color="#66c2a5") + 
            geom_text(x = 45, y = 75, label="R2 adjusted = 0.08, p = 0.084", color="#66c2a5", size=5) +
  geom_point(aes(y=Fine_Hazel), color="#fc8d62", size=3) +
  geom_point(aes(y=Fine_Chestnut), color="#8da0cb", size=3) +
  ylab("Fine mesh size Litter mass loss (%)") + 
  xlab("Bacterial and fungal biomass (µg/g)")

ggsave("microbial_effect.png", height=8, width=8)

lml_wide %>% ggplot(aes(x=sum_arthropods, y=alder_meso)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide %>% ggplot(aes(x=sum_arthropods, y=hazel_meso)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide %>% ggplot(aes(x=sum_arthropods, y=chestnut_meso)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p

lml_wide %>% ggplot(aes(x=sum_arthropods)) + 
  geom_hline(yintercept=0) +
  geom_point(aes(y=alder_meso), color="#66c2a5", size=3) + stat_smooth(aes(y=alder_meso),method="lm", se=FALSE, color="#66c2a5") + 
            geom_text(x = 2400, y = 35, label="R2 adjusted = 0.21, p = 0.056", color="#66c2a5", size=5) +
  geom_point(aes(y=hazel_meso), color="#fc8d62", size=3) + stat_smooth(aes(y=hazel_meso),method="lm", se=FALSE, color="#fc8d62") +
  geom_point(aes(y=chestnut_meso), color="#8da0cb", size=3) + stat_smooth(aes(y=chestnut_meso),method="lm", se=FALSE, color="#8da0cb") +
  ylab("Mesofaunal effect (Δ percent points)") + 
  xlab("Micro-arthropod abundance") 

ggsave("mesofaunal_effect.png", height=8, width=8)

lml_wide %>% ggplot(aes(x=biomass_ew_total, y=alder_macro)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide %>% ggplot(aes(x=biomass_ew_total, y=hazel_macro)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide %>% ggplot(aes(x=biomass_ew_total, y=chestnut_macro)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p

lml_wide %>% ggplot(aes(x=biomass_ew_total)) + 
  geom_hline(yintercept=0) +
  geom_point(aes(y=alder_macro), color="#66c2a5", size=3) +
  geom_point(aes(y=hazel_macro), color="#fc8d62", size=3) + 
  geom_point(aes(y=chestnut_macro), color="#8da0cb", size=3) + geom_smooth(aes(y=chestnut_macro),method="lm", se=FALSE, color="#8da0cb") +
            geom_text(x = 280, y = -28, label="R2 adjusted = 0.26, p = 0.036", color="#8da0cb", size=5) +
  ylab("Macrofaunal effect (Δ percent point)") + 
  xlab("Earthworm biomass") 
  
ggsave("macrofaunal_effect.png", height=8, width=8)


# lml_wide %>% ggplot(aes(x=PC1)) + 
#   geom_point(aes(y=alder_faunal), color="#66c2a5", size=3) + geom_smooth(aes(y=alder_faunal),method="lm", se=FALSE, color="#66c2a5") +
#   geom_point(aes(y=hazel_faunal), color="#fc8d62", size=3) + geom_smooth(aes(y=hazel_faunal),method="lm", se=FALSE, color="#fc8d62") +
#   geom_point(aes(y=chestnut_faunal), color="#8da0cb", size=3) + geom_smooth(aes(y=chestnut_faunal),method="lm", se=FALSE, color="#8da0cb") +
#   ylab("Faunal effect (%)") + 
#   xlab("PC1") 
  
# ggsave("faunal_effect_pc1.png", height=8, width=8)  
# 
# lml_wide %>% ggplot(aes(x=PC2)) + 
#   geom_point(aes(y=alder_faunal), color="#66c2a5", size=3) + geom_smooth(aes(y=alder_faunal),method="lm", se=FALSE, color="#66c2a5") +
#   geom_point(aes(y=hazel_faunal), color="#fc8d62", size=3) + geom_smooth(aes(y=hazel_faunal),method="lm", se=FALSE, color="#fc8d62") +
#   geom_point(aes(y=chestnut_faunal), color="#8da0cb", size=3) + geom_smooth(aes(y=chestnut_faunal),method="lm", se=FALSE, color="#8da0cb") +
#   ylab("Faunal effect (%)") + 
#   xlab("PC2") 
#   
# ggsave("faunal_effect_pc2.png", height=8, width=8)  

```
``` {r faunal effect in percentage}
lml_wide_p <- lml_2 %>% dplyr::select(-c("ff_code","treatment_code","litter_mass_loss","litter_mass_loss_G","notes")) %>% pivot_wider(names_from = c(mesh_size, litter_sp), values_from = litter_mass_loss_final)

lml_wide_p <- lml_wide_p %>% arrange(vb_code)

lml_wide_p <- lml_wide_p %>% mutate(alder_macro_p = ((Coarse_Alder/Medium_Alder)-1)*100,
                    hazel_macro_p = ((Coarse_Hazel/Medium_Hazel)-1)*100,
                    chestnut_macro_p = ((Coarse_Chestnut/Medium_Chestnut)-1)*100,
                    alder_meso_p = ((Medium_Alder/Fine_Alder)-1)*100,
                    hazel_meso_p = ((Medium_Hazel/Fine_Hazel)-1)*100,
                    chestnut_meso_p = ((Medium_Chestnut/Fine_Chestnut)-1)*100)

# filter out only the 4 subplots that have been used for the litter mass experiment
mt_5 <- mt_4 %>% filter(plot_nr < 5)

lml_wide_p <- left_join(lml_wide_p, mt_5, by = "vb_code")

lml_wide_p <- lml_wide_p %>% mutate(sum_arthropods = Mites_oribatid_kg + Mites_other_kg + Springtails_nr_kg,
                    sum_bact_fun = bact_total + Fungi_total)

# PCA only litter mass loss samples
# lml_wide <- lml_wide %>% filter(!grepl("VB002|VB116|VB118",vb_code)) #only use this if you want to plot the faunal effects against PC1/PC2

# lml_wide_long <- lml_wide %>% pivot_longer(cols = c("alder_macro","hazel_macro","chestnut_macro","alder_meso","hazel_meso","chestnut_meso"), values_to="faunal_effect")

lml_wide_p %>% ggplot(aes(x=sum_arthropods, y=alder_meso_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide_p %>% ggplot(aes(x=sum_arthropods, y=hazel_meso_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide_p %>% ggplot(aes(x=sum_arthropods, y=chestnut_meso_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p

lml_wide_p %>% ggplot(aes(x=sum_arthropods)) + 
  geom_hline(yintercept=0) +
  geom_point(aes(y=alder_meso_p), color="#66c2a5", size=3) + stat_smooth(aes(y=alder_meso_p),method="lm", se=FALSE, color="#66c2a5") + 
            geom_text(x = 2400, y = 50, label="R2 adjusted = 0.15, p = 0.091", color="#66c2a5", size=5) +
  geom_point(aes(y=hazel_meso_p), color="#fc8d62", size=3) + stat_smooth(aes(y=hazel_meso_p),method="lm", se=FALSE, color="#fc8d62") +
  geom_point(aes(y=chestnut_meso_p), color="#8da0cb", size=3) + stat_smooth(aes(y=chestnut_meso_p),method="lm", se=FALSE, color="#8da0cb") +
  ylab("Mesofaunal effect (%)") + 
  xlab("Micro-arthropod abundance") +
  ylim(-75,125)

ggsave("mesofaunal_effect_p.png", height=8, width=8)

lml_wide_p %>% ggplot(aes(x=biomass_ew_total, y=alder_macro_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide_p %>% ggplot(aes(x=biomass_ew_total, y=hazel_macro_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide_p %>% ggplot(aes(x=biomass_ew_total, y=chestnut_macro_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p

lml_wide_p %>% ggplot(aes(x=earthworms_2022, y=alder_macro_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide_p %>% ggplot(aes(x=earthworms_2022, y=hazel_macro_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p
lml_wide_p %>% ggplot(aes(x=earthworms_2022, y=chestnut_macro_p)) + geom_point() + stat_poly_line() + stat_poly_eq(use_label(c("adj.R2","p"))) # to get R2 and p

lml_wide_p %>% ggplot(aes(x=biomass_ew_total)) + 
  geom_hline(yintercept=0) +
  geom_point(aes(y=alder_macro_p), color="#66c2a5", size=3) +
  geom_point(aes(y=hazel_macro_p), color="#fc8d62", size=3) +
  geom_point(aes(y=chestnut_macro_p), color="#8da0cb", size=3) +
  ylab("Macrofaunal effect (%)") + 
  xlab("Earthworm biomass") +
  ylim(-100,350)
  
ggsave("macrofaunal_effect_p.png", height=8, width=8)

```

